Estimating individual subjective values of emotion regulation strategies

### 2023-03-03 06_Paper/ER-ED/ERED_Stage2/Manuscript_ERED_Stage2.tex
knitted document

To investigate whether ER strategies require cognitive effort, we conducted an rmANOVA comparing the subjective effort ratings of four strategies (active viewing, distraction, distancing, suppression).
We found a significant effect of strategy (\(F(2.92, 347.65) = 128.47\), \(p < .001\), \(\hat{\eta}^2_G = .327\), 90\% CI \([.261, .384]\), \(\mathrm{BF}_{\textrm{10}} = 1.84 \times 10^{37}\)).
Post-hoc test showed significant higher subjective effort for distraction (\(\eta_{p}^{2}=0.47, [0.42, 1.00]\), \(BF_{10}=\) \(3.61 \times 10^{30}\)), distancing (\(\eta_{p}^{2}=0.41, [0.35, 1.00]\), \(BF_{10}=\) \(1.60 \times 10^{28}\)), and suppression (\(\eta_{p}^{2}=0.30, [0.23, 1.00]\), \(BF_{10}=\) \(1.27 \times 10^{19}\)) compared to active viewing.
Moreover, we found significantly lower effort during suppression compared with distraction (\(\eta_{p}^{2}=0.08, [0.04, 1.00]\), \(BF_{10}=\) \(1.61 \times 10^{6}\)) and distancing (\(\eta_{p}^{2}=0.03, [0.01, 1.00]\), \(BF_{10}=\) 29.19).

@@

To investigate, which variables can predict individual SVs of ER strategies, a multilevel model approach was chosen.
The ICC of the null model was \(ICC=\) 0.19, indicating that level-2 predictor subject accounted for 19.10 percent of total variance.
The preregistered model showed a correlation of \(r=\) 0.95 of the random effects subjects and recoded strategy.
In addition with the estimation that our model explained 0.90 percent of variance, we assumed our model was overfitted due to inclueding recoded strategy as random slope.
We therefore set a new model without the recoded strategy as random slope factor, that followed the specification:

\[
\begin{split}
SV \sim \text{effort rating} + \text{arousal rating} + \text{utility rating} + corrugator \text{ activity} \\\ + levator \text{ activity} + (1 |subject)
\end{split}
\]

@@

\begin{tabular}{lllll}
\toprule
Parameter & \multicolumn{1}{c}{Beta} & \multicolumn{1}{c}{$SE$} & \multicolumn{1}{c}{$p$-value} & \multicolumn{1}{c}{Random Effects (SD)}\\
\midrule
Intercept & 0.80 & 0.01 & 0.00 & 0.12\\
Strategy & -0.22 & 0.01 & 0.00 & 0.14\\
Effort & 0.00 & 0.00 & 0.00 & \\
Arousal & 0.00 & 0.00 & 0.00 & \\
Utility & 0.00 & 0.00 & 0.00 & \\
Corrugator activity & 0.00 & 0.00 & 0.17 & \\
Levator activity & 0.00 & 0.00 & 0.90 & \\
\bottomrule
\end{tabular}

@@

@@

@@

@@

\begin{tabular}{lllllllll}
\toprule
Contrast & \multicolumn{1}{c}{Estimate} & \multicolumn{1}{c}{$SE$} & \multicolumn{1}{c}{$df$} & \multicolumn{1}{c}{$t$} & \multicolumn{1}{c}{$p$} & \multicolumn{1}{c}{$BF10$} & \multicolumn{1}{c}{$\eta_{p}^{2}$} & \multicolumn{1}{c}{$95\% CI$}\\
\midrule
$View_{neg} - Distraction$ & 0.18 & 0.04 & 351.00 & 4.79 & 0.00 & 21,919.73 & 0.06 & {}[0.03, 1.00]\\
$View_{neg} - Distancing$ & 0.19 & 0.04 & 351.00 & 5.09 & 0.00 & 139,814.01 & 0.07 & {}[0.03, 1.00]\\
$View_{neg} - Suppression$ & 0.21 & 0.04 & 351.00 & 5.67 & 0.00 & 18,426,055.49 & 0.08 & {}[0.04, 1.00]\\
$Distraction - Distancing$ & 0.01 & 0.04 & 351.00 & 0.30 & 1.00 & 0.04 & 2.61e-04 & {}[0.00, 1.00]\\
$Distraction - Suppression$ & 0.03 & 0.04 & 351.00 & 0.88 & 1.00 & 0.08 & 2.21e-03 & {}[0.00, 1.00]\\
$Distancing - Suppression$ & 0.02 & 0.04 & 351.00 & 0.58 & 1.00 & 0.05 & 9.51e-04 & {}[0.00, 1.00]\\
\bottomrule
\end{tabular}

@@

\begin{tabular}{lllllllll}
\toprule
Contrast & \multicolumn{1}{c}{Estimate} & \multicolumn{1}{c}{$SE$} & \multicolumn{1}{c}{$df$} & \multicolumn{1}{c}{$t$} & \multicolumn{1}{c}{$p$} & \multicolumn{1}{c}{$BF10$} & \multicolumn{1}{c}{$\eta_{p}^{2}$} & \multicolumn{1}{c}{$95\% CI$}\\
\midrule
$View_{neg} - Distraction$ & 0.34 & 0.05 & 351.00 & 6.73 & 0.00 & 202,428,623,285.88 & 0.11 & {}[0.07, 1.00]\\
$View_{neg} - Distancing$ & 0.28 & 0.05 & 351.00 & 5.66 & 0.00 & 39,887,367.63 & 0.08 & {}[0.04, 1.00]\\
$View_{neg} - Suppression$ & 0.32 & 0.05 & 351.00 & 6.37 & 0.00 & 86,036,487,908.69 & 0.10 & {}[0.06, 1.00]\\
$Distraction - Distancing$ & -0.05 & 0.05 & 351.00 & -1.07 & 1.00 & 0.22 & 3.26e-03 & {}[0.00, 1.00]\\
$Distraction - Suppression$ & -0.02 & 0.05 & 351.00 & -0.36 & 1.00 & 0.04 & 3.70e-04 & {}[0.00, 1.00]\\
$Distancing - Suppression$ & 0.04 & 0.05 & 351.00 & 0.71 & 1.00 & 0.10 & 1.44e-03 & {}[0.00, 1.00]\\
\bottomrule
\end{tabular}

@@

>>>>

To investigate whether ER strategies require cognitive effort, we conducted an rmANOVA comparing the subjective effort ratings of four strategies (active viewing, distraction, distancing, suppression).
We found a significant effect of strategy (\(F(2.92, 347.65) = 128.47\), \(p < .001\), \(\hat{\eta}^2_G = .327\), 90\% CI \([.261, .384]\), \(\mathrm{BF}_{\textrm{10}} = 1.84 \times 10^{37}\)).
Post-hoc test showed significant higher subjective effort for distraction (\(t(357) = -17.92\), \(p_\mathrm{\scriptsize Tukey(4)} < .001\), \(\mathrm{BF}_{\textrm{10}} = 3.61 \times 10^{30}\)), distancing (\(t(357) = -15.82\), \(p_\mathrm{\scriptsize Tukey(4)} < .001\), \(\mathrm{BF}_{\textrm{10}} = 1.60 \times 10^{28}\)), and suppression (\(t(357) = -12.26\), \(p_\mathrm{\scriptsize Tukey(4)} < .001\), \(\mathrm{BF}_{\textrm{10}} = 1.27 \times 10^{19}\)) compared to active viewing.
Moreover, we found significantly lower effort during suppression compared with distraction (\(t(357) = 5.66\), \(p_\mathrm{\scriptsize Tukey(4)} < .001\), \(\mathrm{BF}_{\textrm{10}} = 1.61 \times 10^{6}\)) and distancing (\(t(357) = 3.55\), \(p_\mathrm{\scriptsize Tukey(4)} = .002\), \(\mathrm{BF}_{\textrm{10}} = 29.19\)).

@@

To investigate, which variables can predict individual SVs of ER strategies, a multilevel model approach was chosen.
The ICC of the null model was \(ICC=\) 0.19, indicating that level-2 predictor subject accounted for 19.10 percent of total variance.
The preregistered model showed a correlation of \(r=\) 0.95 of the random effects subjects and recoded strategy (\(BF10\) of the variable strategy: \(\mathrm{BF}_{\textrm{10}} = \infty\)).
Our model explained 0.90 percent of variance and thus we assumed our model was overfitted due to including recoded strategy as random slope.
We therefore set a new model without the recoded strategy as random slope factor to estimate the influence of predictors on SVs more precisely.
The second model followed the specification:
\[
\begin{split}
SV \sim \text{effort rating} + \text{arousal rating} + \text{utility rating} + corrugator \text{ activity} \\\ + levator \text{ activity} + (1 |subject)
\end{split}
\]

The second model explained 0.41 percent of variance.
All results of the second model are in table~@ref(tab:TabMLM\_H5).

@@

\begin{tabular}{llllll}
\toprule
Parameter & \multicolumn{1}{c}{Beta} & \multicolumn{1}{c}{$SE$} & \multicolumn{1}{c}{$p$-value} & \multicolumn{1}{c}{$f^{2}$} & \multicolumn{1}{c}{Random Effects (SD)}\\
\midrule
Intercept & $8.03 \times 10^{-1}$ & 0.01 & 0.00 &  & 0.11\\
Effort & $-6.85 \times 10^{-4}$ & 0.00 & 0.00 & 0.03 & \\
Arousal & $-7.84 \times 10^{-5}$ & 0.00 & 0.32 & 0.00 & \\
Utility & $1.42 \times 10^{-3}$ & 0.00 & 0.00 & 0.15 & \\
Corrugator activity & $7.45 \times 10^{-3}$ & 0.00 & 0.04 & 0.00 & \\
Levator activity & $5.32 \times 10^{-3}$ & 0.00 & 0.07 & 0.00 & \\
\bottomrule
\end{tabular}

@@

The predictors effort rating (\(\hat{\beta} = 0.00\), 95\% CI \([0.00, 0.00]\), \(t(5,618.96) = -13.98\), \(p < .001\)), utility rating (\(\hat{\beta} = 0.00\), 95\% CI \([0.00, 0.00]\), \(t(5,618.96) = -13.98\), \(p < .001\)), and \emph{Corrugator} activity (\(\hat{\beta} = 0.01\), 95\% CI \([0.00, 0.01]\), \(t(5,618.96) = 2.09\), \(p = .037\)) showed a significant association with SVs.
The values were relatively small, so the respective effect size \(f^{2}\) was calculated as the explained variance.
Interestingly, utility rating showed the greatest effect size of all predictors (\(f^{2}=\) 0.155), indicating that utility rating explained 15.5 percent of variance in SVs.
Effort rating showed an effect size of \(f^{2}=\) 0.035.
The effect sizes of all other predictors were negligibly small (\(f^{2}< 0.01\)).

@@

We exploratory wanted to investigate the influence of self-control and the investment trait Need for Cognition (NFC) on individual SVs of ER strategies.
The starting point for this was the model reported before.
Only predictors that had previously shown a significant association with SVs were included in the model together with level 2 predictors self-control and NFC.
The third model followed the specification:
\[
\begin{split}
SV \sim \text{effort rating} + \text{utility rating} + corrugator \text{ activity} \\\ \text{self-control} + \text{NFC} + (1 |subject)
\end{split}
\]
As one would expect, the predictors effort rating (\(\hat{\beta} = 0.00\), 95\% CI \([0.00, 0.00]\), \(t(5,620.93) = -14.26\), \(p < .001\)), utility rating (\(\hat{\beta} = 0.00\), 95\% CI \([0.00, 0.00]\), \(t(5,620.93) = -14.26\), \(p < .001\)), and \emph{Corrugator} activity (\(\hat{\beta} = 0.01\), 95\% CI \([0.00, 0.01]\), \(t(5,620.93) = 2.12\), \(p = .034\)) showed a significant association with SVs.
In addition, a positive association was also found between self-control and SVs (\(\hat{\beta} = 0.02\), 95\% CI \([0.00, 0.05]\), \(t(97.97) = 2.04\), \(p = .044\)).
However, effect size of self-control was negligibly small (\(f^{2}=\) 0.002).
Detailed information can be found in the \protect\hyperlink{SupplementExploratory}{supplementary material}.

@@

\small{

@@

\begin{tabular}{lllllllll}
\toprule
Contrast & \multicolumn{1}{c}{Estimate} & \multicolumn{1}{c}{$SE$} & \multicolumn{1}{c}{$df$} & \multicolumn{1}{c}{$t$} & \multicolumn{1}{c}{$p$} & \multicolumn{1}{c}{$BF10$} & \multicolumn{1}{c}{$\eta_{p}^{2}$} & \multicolumn{1}{c}{$95\% CI$}\\
\midrule
$View_{neg} - Distraction$ & 0.18 & 0.04 & 351.00 & 4.79 & 0.00 & 21,919.73 & 0.06 & {}[0.03, 1.00]\\
$View_{neg} - Distancing$ & 0.19 & 0.04 & 351.00 & 5.09 & 0.00 & 139,814.01 & 0.07 & {}[0.03, 1.00]\\
$View_{neg} - Suppression$ & 0.21 & 0.04 & 351.00 & 5.67 & 0.00 & $1.84 \times 10^{7}$ & 0.08 & {}[0.04, 1.00]\\
$Distraction - Distancing$ & 0.01 & 0.04 & 351.00 & 0.30 & 1.00 & $3.77 \times 10^{-2}$ & 2.61e-04 & {}[0.00, 1.00]\\
$Distraction - Suppression$ & 0.03 & 0.04 & 351.00 & 0.88 & 1.00 & $8.02 \times 10^{-2}$ & 2.21e-03 & {}[0.00, 1.00]\\
$Distancing - Suppression$ & 0.02 & 0.04 & 351.00 & 0.58 & 1.00 & $4.79 \times 10^{-2}$ & 9.51e-04 & {}[0.00, 1.00]\\
\bottomrule
\end{tabular}

@@

\begin{tabular}{lllllllll}
\toprule
Contrast & \multicolumn{1}{c}{Estimate} & \multicolumn{1}{c}{$SE$} & \multicolumn{1}{c}{$df$} & \multicolumn{1}{c}{$t$} & \multicolumn{1}{c}{$p$} & \multicolumn{1}{c}{$BF10$} & \multicolumn{1}{c}{$\eta_{p}^{2}$} & \multicolumn{1}{c}{$95\% CI$}\\
\midrule
$View_{neg} - Distraction$ & 0.34 & 0.05 & 351.00 & 6.73 & 0.00 & $2.02 \times 10^{11}$ & 0.11 & {}[0.07, 1.00]\\
$View_{neg} - Distancing$ & 0.28 & 0.05 & 351.00 & 5.66 & 0.00 & $3.99 \times 10^{7}$ & 0.08 & {}[0.04, 1.00]\\
$View_{neg} - Suppression$ & 0.32 & 0.05 & 351.00 & 6.37 & 0.00 & $8.60 \times 10^{10}$ & 0.10 & {}[0.06, 1.00]\\
$Distraction - Distancing$ & -0.05 & 0.05 & 351.00 & -1.07 & 1.00 & 0.22 & 3.26e-03 & {}[0.00, 1.00]\\
$Distraction - Suppression$ & -0.02 & 0.05 & 351.00 & -0.36 & 1.00 & $3.91 \times 10^{-2}$ & 3.70e-04 & {}[0.00, 1.00]\\
$Distancing - Suppression$ & 0.04 & 0.05 & 351.00 & 0.71 & 1.00 & $9.86 \times 10^{-2}$ & 1.44e-03 & {}[0.00, 1.00]\\
\bottomrule
\end{tabular}

@@

\hypertarget{exploratory-analysis-association-between-svs-and-self-control-and-nfc}{%
\subsection{Exploratory analysis: Association between SVs and self-control and NFC}\label{exploratory-analysis-association-between-svs-and-self-control-and-nfc}}

\caption{\label{tab:TabExp}Exploratory analysis: Results of MLM predicting SVs of ER strategies with level 2 predictors self-control and NFC.}

\begin{tabular}{llllll}
\toprule
Parameter & \multicolumn{1}{c}{Beta} & \multicolumn{1}{c}{$SE$} & \multicolumn{1}{c}{$p$-value} & \multicolumn{1}{c}{$f^{2}$} & \multicolumn{1}{c}{Random Effects (SD)}\\
\midrule
Intercept & $8.03 \times 10^{-1}$ & 0.01 & 0.00 &  & 0.11\\
Effort & $-6.93 \times 10^{-4}$ & 0.00 & 0.00 & 0.04 & \\
Utility & $1.44 \times 10^{-3}$ & 0.00 & 0.00 & 0.20 & \\
Corrugator activity & $7.54 \times 10^{-3}$ & 0.00 & 0.03 & 0.00 & \\
Self-Control & $2.44 \times 10^{-2}$ & 0.01 & 0.04 & 0.00 & \\
NFC & $7.58 \times 10^{-4}$ & 0.00 & 0.44 & 0.00 & \\
\bottomrule
\end{tabular}
